
\begin{resume}

\section*{1.\hspace{0.75em}基本情况}
朱磊，男，陕西西安人，~1982~年~8~月出生，西安电子科技大学~XX~学院~XX~专业~2008~级硕士研究生。
\section*{2.\hspace{0.75em}教育背景}
\begin{resumelist*}
\resumelistitem 2001.08～~2005.07，西安电子科技大学，本科，专业：电子信息工程
\resumelistitem 2008.08～\hspace{3.5em}，西安电子科技大学，硕士研究生，专业：电磁场与微波技术
\end{resumelist*}

\section*{3.\hspace{0.75em}攻读硕士学位期间的研究成果}
\begin{resumelist}{\hspace{-0.25em}3.1\hspace{0.5em} 发表学术论文}
\resumelistitem LEI ZHU, YUXIANG WU, JIETAO LIU, et al. Color imaging through scattering media based on phase retrieval with triple correlation[J].Optics and Lasers in Engineering, 124 (2020), 105796(SCI: 672CZ, EI: 03187452127)
\resumelistitem Tracking moving targets behind a scattering medium via speckle correlation. C Guo, J Liu, T Wu, L Zhu, X Shao. Applied optics 57 (4), 905-913.
\resumelistitem Imaging through scattering layers exceeding memory effect range by exploiting prior information. C Guo, J Liu, W Li, T Wu, L Zhu, J Wang, G Wang, X Shao. Optics Communications 434, 203-208
\resumelistitem Photoacoustic spectrometric evaluation of soil heavy metal contaminants. L Liu, H Huan, M Zhang, X Shao, B Zhao, X Cui, L Zhu. IEEE Photonics Journal 11 (2), 1-7.
\resumelistitem Programmable multiwavelength achromatic focusing and imaging through scattering media. J Liu, J Wang, W Li, X Sun, L Zhu, C Guo, X Shao. IEEE Photonics Journal 10 (5), 1-11
\resumelistitem Recovering the spectral and spatial information of an object behind a scattering media. L Zhu, JT Liu, L Feng, CF Guo, TF Wu, XP Shao
OSA Continuum 1(2), 553-563.
\resumelistitem Single-shot imaging through scattering medium with a deterministic phase-retrieval algorithm. B Tian, L Zhu, B Liu, J Han.  Optics Communications 506, 127562
\resumelistitem Large field-of-view non-invasive imaging through scattering layers using fluctuating random illumination
\resumelistitem 朱磊，邵晓鹏. 散射成像技术的研究进展[J]. 光学学报, 2020, 40(1): 0111005. (EI:    )
\resumelistitem Color imaging through the scattering media
Zhu, L., Wu, Y., Liu, J., Shao, X. 2019 Conference on Lasers and Electro-Optics, CLEO 2019 - Proceedings, 2019, 8750611
\resumelistitem \ldots
\end{resumelist}

\begin{resumelist}{\hspace{-0.25em}3.2\hspace{0.5em} 申请（授权）专利}
\resumelistitem XXX, XXX, XXX等. 专利名称: 国别,专利号[P]. 出版日期.
\resumelistitem \ldots
\end{resumelist}

\begin{resumelist}{\hspace{-0.25em}3.3\hspace{0.5em} 参与科研项目及获奖}
\resumelistitem XXX项目, 项目名称, 起止时间, 完成情况, 作者贡献。
\resumelistitem XXX, XXX, XXX等. 科研项目名称. 陕西省科技进步三等奖, 获奖日期.
\resumelistitem \ldots
\end{resumelist}
\end{resume}
